---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## Introduction

We all know R is not the place to write for-loops, but some of us really, really like them. This package is inspired by Python's list comprehensions that make it possible to write for-loops in only a few keystrokes. Note that this package makes for-loops faster to write, not faster to run.
    
You can install the most recent version of `yuck` from GitHub:

```{r, eval = FALSE}
devtools::install_github("tpq/yuck")
library(yuck)
```

```{r, echo = FALSE}
library(yuck)
```

## Single loop comprehensions

A list comprehension allows a user to step through a for-loop and save each result in one line of code, similar to what `lapply` does. The `yuck` package introduces the `:=` operator to trigger a list comprehension by computing the expression on the right and saving it to the value on the left.

```{r}
a := for(i in 1:5) i^2
a
```

## Nested loop comprehensions

The `yuck` package allows for nested loop comprehensions.

```{r}
a := for(i in 1:5) for(j in 1:7) i * j
matrix(a, 7, 5)
```

See the vignette for more details.
