---
title: "List Comprehensions in R"
author: "Thomas Quinn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{List Comprehensions in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

We all know R is not the place to write for-loops, but some of us really, really like them. This package is inspired by Python's list comprehensions that make it possible to write for-loops in only a few keystrokes. Note that this package makes for-loops faster to write, not faster to run.
    
You can install the most recent version of `yuck` from GitHub:

```{r, eval = FALSE}
devtools::install_github("tpq/yuck")
library(yuck)
```

```{r, echo = FALSE}
library(yuck)
```

## Single loop comprehensions

A list comprehension allows a user to step through a for-loop and save each result in one line of code, similar to what `lapply` does. The `yuck` package introduces the `:=` operator to trigger a list comprehension by computing the expression on the right and saving it to the value on the left.

```{r}
a := for(i in 1:5) i^2
a
```

List comprehensions work with with arbitrary sequences.

```{r}
index <- seq(1, 9, by = 2)
a := for(i in 1:length(index)) index[i]^2
a := for(i in index) i^2
a
```

List comprehensions work with booleans.

```{r}
a := for(i in 1:5) i %in% 1:4
a
```

List comprehensions work with strings.

```{r}
a := for(i in 1:5) paste("this is how we comprehend", i)
a
```


## Nested loop comprehensions

The `yuck` package allows for nested loop comprehensions.

```{r}
a := for(i in 1:5) for(j in 1:7) (i - 1)^2 + (j - 1)^2
a
```

We can save the results of a nested loop comprehension as a matrix.

```{r}
matrix(a, length(1:7), length(1:5))
```

Nested loop comprehensions work with booleans.

```{r}
a := for(i in c(0, 1, 0, 1)) for(j in c(0, 1)) i & j
a <- matrix(a, 2, 4)
colnames(a) <- as.character(c(0, 1, 0, 1))
rownames(a) <- as.character(c(0, 1))
a
```

Nested loop comprehensions work with strings.

```{r}
a := for(i in c(1, 2, 3)) for(j in c("ok now", "whooo")) paste("yes", i, j)
a
```

## Performance

Loops are slow in R, and so too are list comprehensions. Behind the scenes, `:=` pre-allocates RAM to store the results of a for-loop. This makes list comprehensions less slow, but certainly not fast. By default, `lapply` is actually faster.

```{r}
microbenchmark::microbenchmark(
  a := for(i in 1:100000) i^2,
  a <- lapply(1:100000, function(i) i^2),
  times = 5L
)
```

The `yuck` package tries to return the result as a numeric, logical, or character vector. If it cannot, it returns the results as a list.

```{r}
a := for(i in 1:5) rep("A", i)
a
```

## Known issues

List comprehensions use regular expressions to parse the for-loop and save each result. The user should consider the terms `for`, `in`, `~~~`, and `out9000` as protected, meaning that a list comprehension may fail if these terms are present extraneously in an expression.

```{r, eval = FALSE}
# will fail
a := for(i in c("for ", "fail")) i
a := for(i in c("in )", "fail")) i
```

Although you can have nested loop comprehensions, you cannot have a loop comprehension within a loop comprehension (or a for-loop within a function call).

```{r, eval = FALSE}
# will fail
a := for(i in 1:3) b := for(j in 1:3) i * j
```

You can, however, have an `sapply` within a loop comprehension.

```{r}
a := for(i in 1:3) i * sapply(1:i, sum)
a
```

## Use cases

The examples above are arbitrary and the problems presented have better (i.e., vectorized) solutions. Rather, the advantage of `yuck` comes from its use in combination with others tools. For example, we could use `yuck` to iterate across multiple machine learning parameters, models, and predictions. We can also combine `yuck` with `magrittr` to comprehend pipes.

```{r}
library(e1071)
library(magrittr)
data(iris)
groups <- split(sample(1:nrow(iris)), 1:2)
train <- iris[groups[[1]],]
test <- iris[groups[[2]],]
confs := for(kernel in c("linear", "radial", "polynomial"))
  svm(Species ~ ., train, kernel = kernel) %>%
  predict(test) %>% table(test$Species)
confs
```

I like using `yuck` to measure CPU time and RAM overhead.

```{r}
library(peakRAM)
times := for(i in c(1e5, 1e6, 1e7)) data.frame(i,
  peakRAM(
    function() 1:i,
    1:i,
    1:i + 1:i,
    1:i * 2
  ))[,-3]
times
```
